Distance Sampling in \proglang{R}

### 2016-04-05 paper.tex
manually add "R> " to code, urgh!

\begin{CodeChunk}
\begin{CodeOutput}
Warning: package 'nlme' was built under R version 3.2.4
\end{CodeOutput}
\begin{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
library("Distance")
head(minke)
\end{CodeInput}
\begin{CodeOutput}
  Region.Label  Area Sample.Label Effort distance
1        South 84734            1  86.75     0.10
2        South 84734            1  86.75     0.22
3        South 84734            1  86.75     0.16
4        South 84734            1  86.75     0.78
5        South 84734            1  86.75     0.21
6        South 84734            1  86.75     0.95
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
head(amakihi)
\end{CodeInput}
\begin{CodeOutput}
   survey object distance obs mas has detected
1 July 92      1       40 TJS  50   1        1
2 July 92      2       60 TJS  50   1        1
3 July 92      3       45 TJS  50   1        1
4 July 92      4      100 TJS  50   1        1
5 July 92      5      125 TJS  50   1        1
6 July 92      6      120 TJS  50   1        1
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
minke_hn <- ds(minke, truncation = 1.5)
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 46.872
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.872
\end{CodeOutput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
minke_hrcos <- ds(minke, truncation = 1.5, key = "hr")
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.637
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 50.386
\end{CodeOutput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
minke_unifcos <- ds(minke, truncation = 1.5, key = "unif",
                    adjustment = "cos", order = c(1, 2))
\end{CodeInput}
\begin{CodeOutput}
Fitting uniform key function with cosine(1,2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.268
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs <- ds(amakihi, truncation = 82.5, transect = "point",
                     key = "hr", formula = ~obs)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10778.448
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs_mas <- ds(amakihi, truncation = 82.5, transect = "point",
                         key = "hr", formula = ~obs + mas)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10777.376
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
summary(amakihi_hr_obs)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hn <- ds(amakihi, truncation = 82.5, transect = "point",
                 key = "hn", adjustment = NULL)
\end{CodeInput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10833.841
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hn)
\end{CodeInput}
\begin{CodeOutput}

@@

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.93083  P =  0.003578 
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hr_obs_mas)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
summarize_ds_models(amakihi_hn, amakihi_hr_obs, amakihi_hr_obs_mas)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
summary(minke_hn)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
minke_table <- summary(minke_hn)$dht$individuals$N
minke_table$lcl <- minke_table$ucl <- minke_table$df <- NULL
colnames(minke_table) <- c("Stratum", "$\\hat{N}$", "$\\text{se}(\\hat{N}$)", 
                           "$\\text{CV}(\\hat{N}$)")
\end{CodeInput}
\end{CodeChunk}

>>>>

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("Distance")
R> head(minke)
\end{CodeInput}
\begin{CodeOutput}
  Region.Label  Area Sample.Label Effort distance
1        South 84734            1  86.75     0.10
2        South 84734            1  86.75     0.22
3        South 84734            1  86.75     0.16
4        South 84734            1  86.75     0.78
5        South 84734            1  86.75     0.21
6        South 84734            1  86.75     0.95
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> head(amakihi)
\end{CodeInput}
\begin{CodeOutput}
   survey object distance obs mas has detected
1 July 92      1       40 TJS  50   1        1
2 July 92      2       60 TJS  50   1        1
3 July 92      3       45 TJS  50   1        1
4 July 92      4      100 TJS  50   1        1
5 July 92      5      125 TJS  50   1        1
6 July 92      6      120 TJS  50   1        1
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> minke_hn <- ds(minke, truncation = 1.5)
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 46.872
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.872
\end{CodeOutput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> minke_hrcos <- ds(minke, truncation = 1.5, key = "hr")
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.637
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 50.386
\end{CodeOutput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> minke_unifcos <- ds(minke, truncation = 1.5, key = "unif",
                    adjustment = "cos", order = c(1, 2))
\end{CodeInput}
\begin{CodeOutput}
Fitting uniform key function with cosine(1,2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.268
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> amakihi_hr_obs <- ds(amakihi, truncation = 82.5, transect = "point",
                     key = "hr", formula = ~obs)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10778.448
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> amakihi_hr_obs_mas <- ds(amakihi, truncation = 82.5, transect = "point",
                         key = "hr", formula = ~obs + mas)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10777.376
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(amakihi_hr_obs)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> amakihi_hn <- ds(amakihi, truncation = 82.5, transect = "point",
                 key = "hn", adjustment = NULL)
\end{CodeInput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10833.841
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\begin{CodeInput}
R> gof_ds(amakihi_hn)
\end{CodeInput}
\begin{CodeOutput}

@@

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.93083  P =  0.003578 
\end{CodeOutput}
\begin{CodeInput}
R> gof_ds(amakihi_hr_obs_mas)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summarize_ds_models(amakihi_hn, amakihi_hr_obs, amakihi_hr_obs_mas)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(minke_hn)
\end{CodeInput}
\begin{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> minke_table <- summary(minke_hn)$dht$individuals$N
R> minke_table$lcl <- minke_table$ucl <- minke_table$df <- NULL
R> colnames(minke_table) <- c("Stratum", "$\\hat{N}$", "$\\text{se}(\\hat{N}$)", 
                           "$\\text{CV}(\\hat{N}$)")
\end{CodeInput}
\end{CodeChunk}
